## Code by: Jacob Herschberger
## Date: November 2024
## Email: j.herschberger@ufl.edu
## Project: Dissertation chapter 1
### Install and import the required packages.


```{r setup,warning = FALSE, message = FALSE}
packages<-c("semPlot", # Plots the SEMs
            "DBI",# Used to connect to sqlite databases
            "tidyverse", # Great suite of functions for data wrangling and visualization
            "piecewiseSEM", # Create structural equation models with mixed effects and non-normal distributions.
            "lme4", # A package that works well with peicewise SEMs.
            "RSQLite", # Driver to help connect with database
            "glmmTMB",
            "boot",
            "GGally"
            )

for (package in packages) {
if (!require(package, character.only = TRUE)) {
  install.packages("package_name")
  library(package, character.only = TRUE)
} else {
  library(package, character.only = TRUE)
}}

source("Biplot_function.R")
transform_perc <- function(percentage_vec) {
    # See Cribari-Neto & Zeileis (2010)
    (percentage_vec * (length(percentage_vec) - 1) + 0.5)/length(percentage_vec)
}

```

### Import the data
```{r}
con <-dbConnect(SQLite(), '../Data/Data_all_combined.db')
source("Prep_data_for_all_analysis.R")
```


```{r}

ggpairs(PCs %>% select(contains("_PC1")& contains("Clim"),Latitude))

ggpairs(PCs %>% select(contains("_PC2")& contains("Clim"),Latitude))

ggpairs(PCs %>% select(contains("_PC1")& contains("Soil"),Latitude))

ggpairs(PCs %>% select(contains("_PC2")& contains("Soil"),Latitude))

```

### PCA of the climate variables at the field sites.

```{r}
PCbiplot(Clim_ave,rot_x = -1,font_size = 20) 
```
This results show that most climate variables are correlated with latitude whether it be negative or positive.



### Soil data PCA for the field sites.

```{r}
PCbiplot(Soil,rot_y = -1,rot_x = -1,font_size = 20) 
```


### Plant traits PCA of the 2023 common garden.

```{r}
PCbiplot(Garden[,-1:-3])
```
```{r}
Field_2022_mod<-Field_2022 %>% 
  left_join(PCs %>% select(Pop,Soil_PC1,Soil_PC2,
                           Clim_ave_PC1,
                           Clim_ave_PC2)) %>% 
  rename(.,Clim_PC1="Clim_ave_PC1",
         Clim_PC2="Clim_ave_PC2") %>% 
  mutate(Soil_PC2=-Soil_PC2,
         Clim_PC1=-Clim_PC1)
  
 
  
Field_2023_mod<-Field_2023 %>% 
  left_join(PCs %>% select(Pop,Soil_PC1,Soil_PC2,
                           Clim_ave_PC1,
                           Clim_ave_PC2)) %>% 
  rename(.,Clim_PC1="Clim_ave_PC1",
         Clim_PC2="Clim_ave_PC2")%>% 
  mutate(Soil_PC2=-Soil_PC2,
         Clim_PC1=-Clim_PC1)

Garden_mod<-Garden %>% 
  filter(Treatment=="Cont"|Treatment=="us") %>% 
  select(Pop,herb_p,Trichomes,Spines,SLA,Conc,Col,Corolla_D) %>% 
  left_join(PCs %>% select(Pop,Soil_PC1,Soil_PC2,
                           Clim_ave_PC1,
                           Clim_ave_PC2)) %>% 
  rename(.,Clim_PC1="Clim_ave_PC1",
         Clim_PC2="Clim_ave_PC2")%>% 
  mutate(Soil_PC2=-Soil_PC2,
         Clim_PC1=-Clim_PC1,
         Clim_PC1_sq2=scale(Clim_PC1^2))#%>% 
  #group_by(Pop) %>% 
  #summarize_if(is.numeric,mean,na.rm=T)
```

```{r}
ggpairs(Field_2022_mod %>% select(Latitude,Soil_PC1:Clim_PC2))
```


```{r}
Field_combined<-bind_rows(Field_2023_mod,Field_2022_mod) %>% 
  select(!c(Herbivory)) %>% 
  mutate(Clim_PC1_sq2=scale(Clim_PC1^2))#%>% 
  #group_by(Pop) %>% 
  #summarize_if(is.numeric,mean,na.rm=T)
```

## SEM using the combined data from the field.

```{r, message = FALSE,warning = FALSE, results='hide'}
fit<-psem(
  glmmTMB(herb_p~Clim_PC1+Clim_PC1_sq2+Trichomes+Spines+Conc+SLA+Soil_PC2+(1|Pop),Field_combined),
  glmmTMB(Conc~Clim_PC1+Clim_PC1_sq2+Soil_PC2+(1|Pop),Field_combined),
  glmmTMB(SLA~Clim_PC1+Clim_PC1_sq2+Soil_PC2+(1|Pop),Field_combined),
  glmmTMB(Trichomes~Clim_PC1+Clim_PC1_sq2+Soil_PC2+(1|Pop),Field_combined),
  glmmTMB(Spines~Clim_PC1+Clim_PC1_sq2+Soil_PC2+(1|Pop),Field_combined),
 
  Trichomes %~~% SLA,
  SLA %~~% Conc,
  
  data = Field_combined
  )
psem_out<-summary(fit,standardized=T, fit=T,rsquare=T,conserve=T)
```

### Results of the piecewise SEM

```{r}
print(psem_out)
plot(fit)
```

## SEM of the garden data

```{r, message = FALSE,warning = FALSE, results='hide'}
fit<-psem(
  glmmTMB(herb_p~Clim_PC1+Trichomes+Spines+Conc+SLA+Soil_PC2+(1|Pop),Garden_mod),
  glmmTMB(Conc~Clim_PC1+Soil_PC2+(1|Pop),Garden_mod),
  glmmTMB(SLA~Clim_PC1+Soil_PC2+(1|Pop),Garden_mod),
  glmmTMB(Trichomes~Clim_PC1+Soil_PC2+(1|Pop),Garden_mod),
  glmmTMB(Spines~Clim_PC1+Soil_PC2+(1|Pop),Garden_mod),
 
  Trichomes %~~% SLA,
  SLA %~~% Conc,
  SLA %~~% Spines,
  
  data = Garden_mod
  )
psem_out<-summary(fit,standardized=T, fit=T,rsquare=T,conserve=T)
```

### Results of the piecewise SEM

```{r}
print(psem_out)
plot(fit)
```





```{r}
Garden_mod[,"LOC"]<-0
Field_combined[,"LOC"]<-1

QUEST1_data<-rbind(Garden_mod %>% select(-c(Corolla_D,Col)),
                   Field_combined %>% 
                     select(-c(Plant_ID,Date,Corolla_D:Avoid,Latitude)))
```

```{r message=FALSE, warning=FALSE, results='hide'}
fit<-psem(
  glmmTMB(herb_p~(Clim_PC1+Clim_PC1_sq2+Soil_PC2+Trichomes+Spines+Conc+SLA)*LOC+(1|Pop),QUEST1_data),
  glmmTMB(Conc~(Clim_PC1+Clim_PC1_sq2+Soil_PC2)*LOC+(1|Pop),QUEST1_data),
  glmmTMB(SLA~(Clim_PC1+Clim_PC1_sq2+Soil_PC2)*LOC+(1|Pop),QUEST1_data),
  glmmTMB(Trichomes~(Clim_PC1+Clim_PC1_sq2+Soil_PC2)*LOC+(1|Pop),QUEST1_data),
  glmmTMB(Spines~(Clim_PC1+Clim_PC1_sq2+Soil_PC2)*LOC+(1|Pop),QUEST1_data),
 
  SLA %~~%Trichomes,
  SLA %~~% Conc,
  
  data = QUEST1_data
  )
psem_out<-summary(fit,standardized=T, fit=T,rsquare=T,conserve=T)
```


```{r}
print(psem_out)
coeffs<-psem_out$coefficients
plot(fit)
```



### Question 1 figure: Do soil and climate variables predict plant defense traits across a latitudinal gradient?

